---
title: "Getting and plotting ALRN connectivity data"
author: "Marta Costa"
date: "02/12/2020"
output: html_document
---
This file shows an example of getting connectivity data from ALRNs in the hemibrain and plotting it.
The data can be accessed via neuPrint v1.1, using the [`neuprintr`](https://github.com/natverse/neuprintr) package, or the [`hemibrainr`](https://github.com/natverse/hemibrainr) package. 

If using `neuprintr` please note that there might be some minor discrepancies with what is shown in the paper, due to typing changes or differences.

Installation and and set up instructions for both packages are in the links above.

# Set up
Load the necessary libraries.
```{r setup, include=FALSE}
library(knitr)
library(natverse)
library(tidyverse)
library(hemibrainr)
library(neuprintr)
library(ggpubr)

# defaults for plot theme
theme_set(theme_classic(base_size = 12) + theme(text = element_text(family = "sans"), legend.margin=margin(t = 0, unit='cm')))
```

# Prepare data
We will use some curated information for ALRNs already in `hemibrainr`: bodyids and some metadata information about them.

The dataframes `rn.info` includes the list of ALRN bodyids (excludes any outliers or ALRNs pending fixes) while `hemibrain_glomeruli_summary` includes info on the types and status of the glomeruli in the dataset.

For the analysis, we will only use glomeruli that are whole (fragmented and non-fragmented RNs), so first we define those. 
```{r read_data}
hemibrain_glomeruli_summary %>% 
  filter(missing == 0) %>%
  pull(glomerulus) -> whole_frag_gloms

length(whole_frag_gloms)
```
Then subset the list of ALRN bodyids to only those glomeruli.
Add variable for soma_side.
```{r subset_to_whole_gloms}
rn.info %>%
  filter(glomerulus%in%whole_frag_gloms) %>%
    select(bodyid:type, class, side) -> rn.info.whole

nrow(rn.info.whole)
```
# Plot the neurons for one ALRN type
Plot the VL2p ALRNs with an assigned soma: ipsi- and contralateral.
```{r plot_vl2p_alrns}
rn.info.whole %>%
  filter(glomerulus == "VL2p" & side == "ipsilateral") %>% pull(bodyid) -> vl2p_ipsi
rn.info.whole %>%
  filter(glomerulus == "VL2p" & side == "contralateral") %>% pull(bodyid) -> vl2p_contra

# Read in neuron skeletons
vl2p_sk = neuprint_read_neurons(c( vl2p_ipsi, vl2p_contra))

# read in neuropil meshes
al.r = hemibrain_roi_meshes(rois = "AL(R)")
al.l = hemibrain_roi_meshes(rois = "AL(L)")

nopen3d()
# Set view details
op = list(FOV = 0, userMatrix = structure(c(0.99992561340332, -0.0120719093829393, 
0.00176508724689484, 0, -0.00369510054588318, -0.161776661872864, 
0.98682051897049, 0, -0.0116272661834955, -0.986753582954407, 
-0.161809235811234, 0, 0, 0, 0, 1), .Dim = c(4L, 4L)), scale = c(1, 
1, 1), zoom = 0.644609212875366, windowRect = c(1680L, 45L, 2468L, 
690L))

# plot skeletons showing synapses
plot3d(vl2p_sk[as.character(vl2p_ipsi)], col = hemibrain_bright_colours["yellow"], WithConnectors = T, lwd=2)
plot3d(vl2p_sk[as.character(vl2p_contra)], col = hemibrain_bright_colours["darkorange"], WithConnectors = T, lwd=2)
plot3d(al.r, col = "grey90", alpha=0.1, add=T)
plot3d(al.l, col = "grey90", alpha=0.1, add=T)
par3d(op)
```

# Plot fraction of output per ALRN type
Define what is not an ALRN, PN, ALLN but has been identified.
```{r other_neurons_class}
class_other = c(dan.ids, mbon.ids, ton.ids, lhn.ids, dn.ids, kc.ids, apl.ids, cent.ids, lc.ids)
```

Read in connectivity data downstream of the ALRNs (those in whole glomeruli) and add metadata for the partners.
```{r read_conn_data}
# connectivity data for ALRNs
alrn_outconn = neuprint_connection_table(rn.info.whole$bodyid, prepost = "POST", roi = "AL(R)", all_segments = F)
  
# add class to partner
alrn_outconn %>%
  left_join(rn.info.whole) %>%
  mutate(class_partner = case_when(
    partner%in%rn.ids ~ "ALRN",
    partner%in%pn.ids ~ "PN",
    partner%in%alln.ids ~ "ALLN",
    partner%in%class_other ~ "other",
    TRUE ~ "unknown"))-> alrn_outconn
```
Calculate the connectivity fractions, per ALRN type.
```{r get_fraction}
# Calculate total connectivity per type
alrn_outconn %>%
  count(glomerulus, modality, wt=ROIweight, name = "syntotal") -> alrn_out_glom_syntotal

# Calculate the fraction of connectivity for class of partner
alrn_outconn %>%
  count(glomerulus, modality, class_partner, wt = ROIweight, name = 'syn') %>%
  left_join(alrn_out_glom_syntotal) %>%
  mutate(class_partner = fct_relevel(class_partner, "PN", "ALLN", "ALRN", "other","unknown"),
         synfrt = syn/syntotal) -> alrn_out_glom_summ

# get order of glomeruli
alrn_out_glom_summ %>%
  ungroup() %>%
  filter(class_partner == "PN") %>%
  group_by(modality) %>%
  arrange(desc(synfrt), .by_group = T) %>%
  pull(glomerulus) -> ordered_gloms_out
```
Plotting the fraction of output per partner class per ALRN type.
```{r plot_alrn_output}
alrn_out_glom_summ %>%
  mutate(glomerulus = fct_relevel(glomerulus, ordered_gloms_out)) %>%
  ggplot(aes(x = glomerulus, y=synfrt, fill = fct_rev(class_partner))) +
  geom_bar(position="stack", stat="identity") +
  facet_grid( ~ modality, scales = "free_x", space = "free_x") +
  scale_fill_manual(values = c("PN" = hemibrain_bright_colors[["brown"]], 
                               "ALLN" = hemibrain_bright_colors[["green"]],
                               "ALRN" = hemibrain_bright_colors[["yellow"]], 
                               "other" = "black", 
                               "unknown" = "grey70")) +
  labs(x = "ALRN type", y = "fraction of ALRN output", fill = "from ALRN to:") +
  scale_y_continuous(expand = (c(0.00, 0.01))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 1), 
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.line.y = element_blank(), 
        axis.ticks.y = element_blank(),
        strip.background = element_blank(),
        ) +
  guides(fill = guide_legend(reverse = TRUE))
```
# Compare the fraction of of output per soma side of ALRN
For this, we will only use the glomeruli that are whole and do not include fragmented ALRNs. Since only 1/7 TRNs or HRNs are bilateral, we use only bilateral ORNs.

```{r get_conn_soma_side}
hemibrain_glomeruli_summary %>% 
  filter(missing == 0 & 
           RN_frag == "no" & 
           str_detect(glomerulus, "^VP", negate = TRUE) & 
           laterality == "bilateral") %>%
  pull(glomerulus) -> whole_nofrag_bil_gloms

alrn_outconn %>% 
  filter(glomerulus%in%whole_nofrag_bil_gloms) -> bilorn_outconn
```
Calculate the fraction per type and soma side.
Exclude any ALRNs for which the soma side has not been assigned.
```{r prepare_conn_soma_side}
bilorn_outconn %>%
  filter(!is.na(side)) %>%
  count(side, glomerulus, modality, wt=ROIweight, name = "syntotal") -> bilorn_out_glom_syntotal
  
bilorn_outconn %>%
  filter(!is.na(side)) %>%
  count(side, glomerulus, modality, class_partner, wt = ROIweight, name = 'syn') %>%
  left_join(bilorn_out_glom_syntotal) %>%
  mutate(class_partner = fct_relevel(class_partner, "PN", "ALLN", "ALRN", "other","unknown"),
         side = fct_relevel(side, "ipsilateral", "contralateral"),
         synfrt = syn/syntotal) -> bilorn_outconn_glom_summ
```
Compare means between the soma side, for each class of partner.
```{r compare_means_soma_side}
bilorn_outconn_glom_summ %>%
  filter(class_partner%in%c("PN", "ALLN", "ALRN")) %>%
  select(side, class_partner, synfrt) -> test

compare_means(synfrt ~ side, test, group.by = "class_partner")
```
Make a paired plot.
```{r soma_side_out_paired_plot}
bilorn_outconn_glom_summ %>%
  filter(class_partner%in%c("PN", "ALLN", "ALRN")) %>%
  mutate(class_partner = str_glue("ORN to ", "{class_partner}"),
         class_partner = fct_relevel(class_partner, "ORN to PN", "ORN to ALLN", "ORN to ALRN")) %>%
  ggplot(aes(x = side, y=synfrt)) +
  geom_path(aes(group = glomerulus), col = "grey80", size=0.5) +
  geom_point(col = "grey80", show.legend = FALSE) +
  stat_summary(aes(col = side), fun.data = "mean_sdl", fun.args = list(mult = 1), 
               pch = 18, size = 1.2, show.legend = FALSE) +
  facet_grid( ~ class_partner) +
  stat_compare_means(comparisons = list(c("ipsilateral", "contralateral")), synfrt ~ side, paired = TRUE, label = "p.signif") +
  scale_colour_manual(values = c("ipsilateral" = hemibrain_bright_colors[["yellow"]],
                                 "contralateral" = hemibrain_bright_colors[["orange"]])) +
  scale_x_discrete(labels = c("ipsilateral" = "ipsi", "contralateral" = "contra")) +
  scale_y_continuous(expand = (c(0.00, 0.01)), limits = c(0, 1)) +
  theme(strip.background = element_blank()) +
  labs(x = "ORN soma side", y = "fraction of ORN output", col = NULL)
```
